---
title: Effect of the Pandemic on Azithromycin Prescription in General Practices across the UK
author: "Daramola Oluwasegun Isaac"
date: "8/4/2021"
output:
  html_document: default
  pdf_document: default
---

# Load Packages
```{r run-packages, message=FALSE, warning=FALSE}
# List of required packages
packages <- c("fpp2", "ggplot2", "tidyverse", "pandoc")

# Install any missing packages
installed_packages <- packages %in% rownames(installed.packages())
if(any(!installed_packages)) {
  install.packages(packages[!installed_packages])
}

# Load the installed packages
library(fpp2)
library(ggplot2)
library(tidyverse)

```

# Load Main Sheets (Data for both and Pre-covid)
```{r load-sheets, message=FALSE, warning=FALSE}
precovid <- read.csv("Azidata/Aziprecovid.csv", as.is = T)
both <- read.csv("Azidata/Aziboth.csv", as.is = T)
view(precovid)
view(both)
```

# Convert needed data columns to time series 
```{r convert-to-ts, message=FALSE, warning=FALSE}
pre <- ts(precovid, start = c(2016, 5), frequency = 12)
preazi <- ts(precovid$Azithromycin, start = c(2016, 5), frequency = 12)
patients <- ts(precovid$Patients, start = c(2016, 5), frequency = 12)
bothazi <- ts(both$Azithromycin, start = c(2016, 5), frequency = 12)
```

# Plot Azithromycin prescription
```{r Plot-azithromycin-prescription, message=FALSE, warning=FALSE}
autoplot(preazi, main = "Pre-Covid Azithromycin prescription (UK) time series")
autoplot(bothazi, main = "Azithromycin prescription (UK) time series")
```

# Lag plot and autocorrelation
```{r lag-plots, message=FALSE, warning=FALSE}
preazilag <- window(preazi, start = c(2016, 5))
gglagplot(preazilag)
ggAcf(preazilag, lag = 44)

preazidiff <- diff(preazi)
autoplot(preazidiff, main = "Pre-Covid Azithromycin prescription (UK) monthly differences")
ggAcf(preazidiff, lag = 44)

bothazilag <- window(bothazi, start = c(2016, 5))
gglagplot(bothazilag)
ggAcf(bothazilag, lag = 61)

bothazidiff <- diff(bothazi)
autoplot(bothazidiff, main = "Azithromycin prescription (UK) monthly differences")
ggAcf(bothazidiff, lag = 61)
```

# Box-Cox Transformation
```{r Box-Cox-transformation, message=FALSE, warning=FALSE}

(lambdaprecovid <- BoxCox.lambda(preazilag))
autoplot(BoxCox(preazilag, lambdaprecovid))
```

# Residual checks
```{r residual-checks, message=FALSE, warning=FALSE}
preres <- residuals(naive(preazi))
autoplot(preres) + xlab(" ") + ylab(" ") +
  ggtitle("Residuals from naive method")
gghistogram(preres) + ggtitle("Histogram of residuals")
ggAcf(preres) + ggtitle("ACF of residuals")
```

# Portmanteau Tests
```{r lPortmanteau tests, message=FALSE, warning=FALSE}
Box.test(preres, lag = 10, fitdf = 0)
Box.test(preres, lag = 10, fitdf = 0, type = "Lj")
checkresiduals(naive(preazi))
```

# Basic principle forecasts
```{r simple-forecasts, message=FALSE, warning=FALSE}
autoplot(preazi) +
  autolayer(meanf(preazi, h = 17),
            series = "Mean", PI = FALSE) +
  autolayer(naive(preazi, h = 17),
            series = "Naive", PI = FALSE) +
  autolayer(snaive(preazi, h = 17),
            series = "Seasonal Naive", PI = FALSE) +
  autolayer(snaive(preazi)) +
  ggtitle("Forecasts for Azithromycin Prescription") +
  xlab("Year") + ylab("Azithromycin Prescription") +
  guides(color=guide_legend(title="Forecast"))
```

# Variable Plot Comparison
```{r variable comparison, message=FALSE, warning=FALSE}
autoplot(pre[, c("Azithromycin", "Patientsthousand")])
```

# Correlation Plot
```{r correlation plot, message=FALSE, warning=FALSE}
GGally::ggpairs(as.data.frame(pre[, c("Azithromycin", "Patientsmillion")]))

pre %>% 
  as.data.frame() %>% 
  ggplot(aes(x = Azithromycin, y = Patientsmillion)) +
  ylab("Patients (millions)") +
  xlab("Azithromycin Prescription") +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
tslm(Patientsmillion ~ Azithromycin, data = pre)
```

# Fit regression plot for both variables
```{r variable-regression-plot, message=FALSE, warning=FALSE}
fit.pre <- tslm(
  Azithromycin ~ Patients,
  data=pre
)

autoplot(pre[, "Azithromycin"], series = "Data") +
  autolayer(fitted(fit.pre), series = "Fitted")

cbind(Data = pre[, "Azithromycin"],
      Fitted = fitted(fit.pre)) %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Data, y=Fitted)) +
  geom_point() +
  geom_abline(intercept=0, slope = 1)
checkresiduals(fit.pre)

df <- as.data.frame(pre)
df[, "Residuals"] <- as.numeric(residuals((fit.pre)))
ggplot(df, aes(preazi, Residuals)) +
  geom_point()
cbind(Fitted= fitted(fit.pre),
      Residuals= residuals(fit.pre)) %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Fitted, y=Residuals)) + geom_point()
summary(fit.pre)
CV(fit.pre)
```

# Linear Regression plot
```{r linear-regression-plot, message=FALSE, warning=FALSE}
prefc <- window(preazi, start = c(2016, 5))
fit.prefc <- tslm(preazi ~ trend + season)
fcast <- forecast(fit.prefc, h = 17)
autoplot(fcast) +
  autolayer(bothazi, colour = FALSE) +
  ggtitle("Forecast of Azithromycin prescription using Linear regression") +
  xlab("Year") + ylab(" ")

summary(fit.prefc)
CV(fit.prefc)

preazi %>% decompose(type="multiplicative") %>% 
  autoplot() + xlab("Year")
```

# Export New Prediction data

After run, some off-R modification to calculate percentage change
% Change = ((Actual Value - Predicted Value)/Actual Value) in Excel.
Insert new columns (% change in point forecast, 95% prediction interval high and low)
```{r export-forecast-data, message=FALSE, warning=FALSE}
write.csv(fcast, file = "fc.csv")
```

# Load % change sheet for differences

"fc.csv" file with changes now named "fc2.csv"
```{r Percentage change, message=FALSE, warning=FALSE}
fc <- read.csv("fc2.csv", as.is = T)

percentchange <- ts(fc$PercentageChange, start = 2020, frequency = 12)
percentdecrease <- ts(fc$PercentDecrease, start = 2020, frequency = 12)
percentincrease <- ts(fc$PercentIncrease, start = 2020, frequency = 12)
change <- ts(fc, start = 2020, frequency = 12)

autoplot(percentchange, series = "Point Forecast") +
  autolayer(percentdecrease, series = "95% Prediction Interval (Low)") +
  autolayer(percentincrease, series = "95% Prediction Interval (High)") +
  geom_hline(aes(yintercept = 0))
```